---
title: "projectversion2"
output: html_document
date: "2023-01-19"
---

```{r libraries data, message=FALSE, warning=FALSE}
# Libraries

library(readxl)
library(readr)
library(tidyverse)
library(dplyr)
library(knitr)
library(kableExtra)
library(magrittr)
library(RColorBrewer)


 
#Data sets(add importing codes here)
MyData =  readxl::read_excel("Datasets/M2Data.xlsx")

read_excel("Datasets/M2Data.xlsx")

head(MyData, n=5)

tail(MyData, n=5)

knitr::kable(head(MyData[, 1:2],"simple"))

Market = MyData$Market

table(MyData$Market)
as.data.frame(table(MyData$Market))

Market %>%
  table() %>%
  barplot()

Mafrica = filter(MyData, Market == 'Africa')
 filter(Mafrica, Market == 'Africa')
 
 t4Africa = dplyr::filter(Mafrica, Market == 'Africa')
knitr::kable(head(Mafrica[], 'simple'))

task_p = Mafrica$Product_Category
table(Mafrica$Product_Category)
as.data.frame(table(task_p))

task_p %>%
  table(Mafrica$Product_Category) %>%
  pie()
 
```{r}
task_t = Mafrica$Product_SubCategory
table(Mafrica$Product_SubCategory)
as.data.frame(table(task_t))

```{r}
#Create a table
task5_table = table(Mafrica$Product_SubCategory)

#Plot the table using a bar plot
t5bar = barplot(task5_table)

#use text() to show values per column
text(y=table(Mafrica$Product_SubCategory),
     t5bar,
     table(Mafrica$Product_SubCategory),
     cex=0.8,
     pos=3)
#add some par codes
```{r}

task_q = Mafrica$Sales

tapply(task_q,task_t, mean)
as.data.frame(tapply(task_q,task_t, mean))

#create object 'meanSales' to store data
meanPetal = tapply(iris$Petal.Length, iris$Species, mean)
#present data using table
knitr::kable(meanPetal)

dotchart(meanPetal)

```{r}
task_r = Mafrica$Region
table(Mafrica$Region)
as.data.frame(table(task_r))

```{r}
#Create a table
task8_table = table(Mafrica$Region)

#Plot the table using a bar plot
t8bar = barplot(task8_table)

#use text() to show values per column
text(y=table(Mafrica$Region),
     t8bar,
     table(Mafrica$Region),
      cex=0.8,
     pos=3)



```{r}
task_q = Mafrica$Sales
tapply(task_q, task_r, sum)
as.data.frame(tapply(task_q, task_r, sum))


task_s = Mafrica$ShippingCost
tapply(task_s, task_r, mean)
as.data.frame(tapply(task_s, task_r, mean))

```{r}
#Create a table
task9_table = table(Mafrica$Region)

#Plot the table using a bar plot
t9bar = barplot(task9_table)

#use text() to show values per column
text(y=table(Mafrica$Region),
     t9bar,
     table(Mafrica$Region),
     cex=0.8,
     pos=3)

```{r}

###task 11
par(mfcol=c(2,1),
    mai=c(1, 1, 0.2, 0.4),
    mar=c(4,4,0.5,2))

#present a box plot of all profits
boxplot(MyData$Profits,
        horizontal = T)

#present a histogram of all profits
hist(MyData$Profits,
     breaks = 50,
     main = "")
```{r}

t13LATAM = filter(MyData, Market == 'LATAM')
 filter(t13LATAM, Market == 'LATAM')
 knitr::kable(head(t13LATAM[], 'simple'))
 
task_a = t13LATAM$Region
table(t13LATAM$Region)
as.data.frame(table(task_a))

#Create a table
task14_table = table(t13LATAM$Region)



                                        # Create example data
t14bar = data.frame('South_America' = rnorm(1000),
                   'Caribbean' = runif(1000),
                   'Central_America' = rpois(1000, 3))
head(t14bar)

#create a boxplot
boxplot(t14bar)

```{r}
t15 = MyData$Product_Category %>%
  table() %>%
  as.data.frame() %>%
  rename(Frequency = Freq) %>%
  mutate(Cumsum = cumsum(Frequency), 
         Probabilities = Frequency/nrow(MyData),
         Cumprob = cumsum(Probabilities))
knitr::kable(t15,
             digits = 2,
             caption = "Table's name to be fixed",
             format = "html",
             table.attr = "style ='width:40%;'",
             align = 'c') %>%
  kable_classic(bootstrap_options = "striped",
                full_width = TRUE,
                position = "center",
                font_size = 12)


 t15 = MyData$Product_Category %>%
   table()%>%
  pie()



  


 
 
 





